Viscous Erosion in Porous Media

### 2018-02-26 docs/methods/erosion_methods.tex
NM

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.75 \textwidth]{./figs/shrink_intface.pdf}
\caption{
Shape and stress evolution: (a) Interfaces of the eroding body from Fig.~\ref{01bodseq} at evenly-spaced time intervals. Color indicates normalized time, $t/t_f$, where $t_f = 1.79E-2$ is the time at which the body completely vanishes. (b) The distribution of absolute shear stress, $\atau$, along each of these interfaces. Normalized arc length, $s/L$, runs counterclockwise from the rear stagnation point. The stress initially varies over the body's surface, but becomes more uniform over time.}
\label{shrink_intface}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0

@@

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). Certain features of this shape can be predicted analytically by appealing to the principal that, in long time, the shear stress becomes uniformly distributed along the body. In particular, consider the angle formed at the front and rear stagnation points. In a small region around this point, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge. We aim to determine the wedge's opening angle, $\oangle$, through the principle of uniform stress. Exact solutions can be obtained to this problem via separation of variables in polar coordinates $(r, \theta)$~\cite{poz1997}. In this section alone, $\theta$ denotes the polar coordinate (not the local tangent angle along the body). The corresponding stream function is assumed to take the form $\psi = r^{\lambda}f(\theta)$, with $\Delta^2 \psi = 0$, which gives the 4th order ODE~\cite{poz1997}
\begin{align*}
  f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0
\end{align*}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$, where $\thb = \pi - \oangle/2$, along with the condition of odd-symmetry $f(0) = f''(0) = 0$ (corresponding to up-down symmetry of the flow).  The corresponding shear stress evaluated on the surface of the wedge is $\tau = \mu r^{\lambda-2} f''(\thb)$. Thus, the wedge of uniform stress corresponds to $\lambda = 2$. For this value, the solution for $f$ satisfying odd symmetry is
\begin{align*}
  f(\theta) = B \sin (2 \theta) + C \theta
\end{align*}
Imposing the no-slip boundary conditions gives the condition
\begin{align*}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0
\end{align*}
which has a solution $\thb \approx 129$ degrees. This gives an opening angle of $\oangle = 2(180^{\circ} - \thb) \approx 102$ degrees. 
% Pozrikidis page 239: symmetric flow.

>>>>

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{./figs/shrink_intface.pdf}
\caption{
Shape and stress evolution: (a) Interfaces of the eroding body from Fig.~\ref{01bodseq} at evenly-spaced time intervals. Color indicates normalized time, $t/t_f$, where $t_f = 1.79E-2$ is the time at which the body completely vanishes. (b) The distribution of absolute shear stress, $\atau$, along each of these interfaces. Normalized arc length, $s/L$, runs counterclockwise from the rear stagnation point. The stress initially varies over the body's surface, but becomes more uniform over time.}
\label{shrink_intface}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0

@@

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). Certain features of this shape can be predicted analytically by appealing to the principal that, in long time, the shear stress becomes uniformly distributed along the body. In particular, consider the angle formed at the front and rear stagnation points. In a small region around this point, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge. We aim to determine the wedge's opening angle, $\oangle$, through the principle of uniform stress. Exact solutions can be obtained to this problem via separation of variables in polar coordinates $(r, \theta)$~\cite{poz1997}. In this section alone, $\theta$ denotes the polar coordinate (not the local tangent angle along the body). The corresponding stream function is assumed to take the form $\psi = r^{\lambda}f(\theta)$, with $\Delta^2 \psi = 0$, which gives the 4th order ODE~\cite{poz1997}
\begin{align*}
  f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0
\end{align*}

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.4 \textwidth]{./figs/corner.pdf}
\caption{2D Stokes flow near a corner: The flow around an infinite wedge of opening angle $\oangle$ is described by a stream-function $\psi(r,\theta)$ for $\theta \in [-\theta_0, \theta_0]$}.
\label{corner}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$, where $\thb = \pi - \oangle/2$, along with the condition of odd-symmetry $f(0) = f''(0) = 0$ (corresponding to up-down symmetry of the flow).  The corresponding shear stress evaluated on the surface of the wedge is $\tau = \mu r^{\lambda-2} f''(\thb)$. Thus, the wedge of uniform stress corresponds to $\lambda = 2$. For this value, the solution for $f$ satisfying odd symmetry is
\begin{align*}
  f(\theta) = B \sin (2 \theta) + C \theta
\end{align*}
Imposing the no-slip boundary conditions gives the condition
\begin{align*}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0
\end{align*}
which has a solution $\thb \approx 129$ degrees. This gives an opening angle of $\oangle = 2(180^{\circ} - \thb) \approx 102$ degrees. 
% Pozrikidis page 239: symmetric flow.
